---
title: "Reproductible Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Course Project 1

## Introduction

Extract from Coursera Website and consignes.

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

<https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip>

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

date: The date on which the measurement was taken in YYYY-MM-DD format

interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Submission 

The library are charged 

````{r}
library("ggplot2")
library("dplyr")
```

Then the data are readed.

## 1. Code for reading in the dataset and/or processing the data.

```{r}
data <- read.csv("activity.csv", sep = ",", colClasses = c("numeric", "character", "numeric"))
```

The data are grouped by day and then summarise by the number of steps the NA are removed. 

```{r}
dataDay <- group_by(data, date)
FinalData <- summarise(dataDay, totalSteps = sum(steps, na.rm = TRUE))
```

## 2. Histogram of the total number of steps taken each day

```{r}
ggplot(FinalData, aes(x=totalSteps)) + geom_histogram()
```

## 3. Mean and median number of steps taken each day.

First the mean. 

```{r}
mean(FinalData$totalSteps)
```

Then the median

```{r}
median(FinalData$totalSteps)
```

## 4. Time series plot of the average number of steps taken.

The original data are taken, and this time group by the interval, then the data are summerized to make the average by interval and then plot.

```{r}
dataCout <- group_by(data, interval)
countPlot <- summarise(dataCout, avgSteps = mean(steps, na.rm = TRUE))
plot(countPlot , type= "l")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps.

First the interval with the maximum average number of steps is search, then the interval is extracted and read. 

Not a very nice or clean way to do it. Sorry. 

```{r}
max(countPlot$avgSteps)
nbmax <- subset(countPlot, countPlot$avgSteps > 206)
nbmax$interval
```

## 6. Code to describe and show a strategy for imputing missing data.

To deal with the missing value. I choose to give them the average value of the interval without the missing value. It is a classical way to complete with the missing value. And other way would be to complete with the neighbour value. The value before or after the missing value. But after a close look to the original data there is often a branch of value that are missing therefore I think conpute the average are a better solution. 

First I made a copy of the original data to. 

```{r}
data2 <- data
```

Then a for loop is used to search the missing value and then the average is exctracted from the previous dataset. Finally the average value is computed insted of the missing value. 

```{r}
for( i in 1:17568){
  if(is.na(data2[i,1])){
    inter <- data[i,3]
    newData <- dataCout[,2] == inter
    data2[i,1] <- newData[1,1]
  }
}
```

After the data is completed, the transformation are applied as for the first data set to plot them. 

```{r}
dataDay2 <- group_by(data2, date)
FinalData2 <- summarise(dataDay2, totalSteps = sum(steps, na.rm = TRUE))
```


## 7. Histogram of the total number of steps taken each day after missing values are imputed.

The data are grouped by date and summerise to ave the total number of step. They can be plot. 

```{r}
ggplot(FinalData2, aes(x=totalSteps)) + geom_histogram()
```


## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

In a first step the date are extracted from the data. It may have been better to do it from the begining. Then the vectore of type character is transform in a vectore of type date. And the format are change so the date will be write with the name of the day. Then the name of the day are combine with the data set. 

```{r}
date <- data[,2]
date2 <- as.Date(date)
Weekday <- weekdays(date2)
data3 <- cbind(data, Weekday)
```
Two dataset are created one for the week day and another for the weekend day. Both are then group by interval and then summarise to have the average number of steps by interval.

```{r}
workDay <- filter(data3, Weekday == "lundi" | Weekday == "mardi" | Weekday == "mercredi" | Weekday == "jeudi" 
                  | Weekday == "vendredi")
weekEnd <- filter(data3, Weekday == "samedi" | Weekday == "dimanche")

workDayCount <- group_by(workDay, interval)
workDayPlot <- summarise(workDayCount, avgSteps = mean(steps, na.rm = TRUE))

weekEndCount <- group_by(weekEnd, interval)
weekEndPlot <- summarise(weekEndCount, avgSteps = mean(steps, na.rm = TRUE))
```
With this two data, the two plot can be made. Of the avergae number of steps make during the weekend and during the week.

```{r}
par(mfrow = c(1,2), mar = c(5,4,3,1))
plot(workDayPlot, main = "Average number of Steps \n on Work Days" , xlab = "Interval" , ylab = "Avg nb of steps",
     pch = 20, type = "l")
plot(weekEndPlot, main = "Average number of Steps \n on Weekend", xlab = "Interval" , ylab = "Avg nb of steps",
     pch = 20, type = "l")
```

